Lapse & Vision Loss PS Analysis

### 2025-02-24 analysis.Rmd
wip

library(readr)
library(dplyr)
library(kableExtra)
library(purrr)
library(gee)
library(geepack)
library(ggplot2)
library(MatchIt)
library(twang)
library(survey)
library(cobalt)
library(margins)
library(margins)

@@

# Define function to run matching with disk storage and selective re-running
run_matching_methods <- function(data, formula, methods, rerun_all = FALSE, output_dir = "ps_results") {
  # Ensure the specified output directory exists
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }

@@

  # Iterate over each method
  for (params in methods) {
    filename <- file.path(output_dir, paste0(params$name, ".rds"))

@@

    if (file.exists(filename) && !params$rerun && !rerun_all) {
      print("Results are cached. Proceeding to next method...")
      method_results <- readRDS(filename)
      all_results[[params$name]] <- method_results
      next
    }

@@

    # For Twang
    if (params$method == "twang") {
      result <- tryCatch({
        ps(formula = formula, data = data.frame(data), 
           n.trees = 10000,
           interaction.depth = 3,
           shrinkage = 0.01,
           estimand = params$estimand,
           stop.method = c("es.mean", "ks.max"))  # Stopping criteria
      }, error = function(e) {
        message("Error in TWANG GBM: ", e)
        return(NULL)
      })

      if (!is.null(result)) {
        balance_plot <- tryCatch({
          love.plot(result, stat = "mean.diffs", var.order = "es.mean", stars="std")
        }, error = function(e) {
          message("Error generating TWANG balance plot: ", e)
          NULL
        })

        common_support_plot <- tryCatch({
          bal.plot(result, var.name = "prop.score.es.mean", which="both")
        }, error = function(e) {
          message("Error generating TWANG common support plot: ", e)
          NULL
        })

        balance_summary <- tryCatch({
          summary(result)
        }, error = function(e) {
          message("Error generating TWANG balance summary: ", e)
          NULL
        })
      }

    } 
    # For WeightIt
    else if(length(params$use) > 0 && params$use == "weightit") {
       result <- tryCatch({
        do.call(weightit, c(list(formula = formula, data = data), params))
      }, error = function(e) {
        message("Error in ", params$name, ": ", e)
        return(NULL)
      })

      if (is.null(result)) next

      balance_plot <- tryCatch({
        love.plot(result, stat = "mean.diffs", var.order = "adjusted", stars="std")
      }, error = function(e) {
        message("Error generating balance plot for ", params$name, ": ", e)
        NULL
      })

      if(! params$method %in% c("nearest_mahalanobis", "cem", "cbps", "entropy")) {
        common_support_plot <- tryCatch({
          bal.plot(result, var.name = "distance", which = "both")
        }, error = function(e) {
          message("Error generating common support plot for ", params$name, ": ", e)
          NULL
        })
      } else {
        common_support_plot <- NULL
      }

      balance_summary <- tryCatch({
        bal.tab(result)
      }, error = function(e) {
        message("Error generating balance summary for ", params$name, ": ", e)
        NULL
      })
    }
    # For MatchIt
    else {
      result <- tryCatch({
        do.call(matchit, c(list(formula = formula, data = data), params))
      }, error = function(e) {
        message("Error in ", params$name, ": ", e)
        return(NULL)
      })

      if (is.null(result)) next

      balance_plot <- tryCatch({
        love.plot(result, stat = "mean.diffs", var.order = "adjusted", stars="std")
      }, error = function(e) {
        message("Error generating balance plot for ", params$name, ": ", e)
        NULL
      })

      if(! params$method %in% c("nearest_mahalanobis", "cem")) {
        common_support_plot <- tryCatch({
          bal.plot(result, var.name = "distance", which = "both")
        }, error = function(e) {
          message("Error generating common support plot for ", params$name, ": ", e)
          NULL
        })
      } else {
        common_support_plot <- NULL
      }

      balance_summary <- tryCatch({
        summary(result)
      }, error = function(e) {
        message("Error generating balance summary for ", params$name, ": ", e)
        NULL
      })
    }

    method_results <- list(
      result = result,
      balance_plot = balance_plot,
      common_support_plot = common_support_plot,
      summary = balance_summary
    )

    saveRDS(method_results, file = filename)
    all_results[[params$name]] <- method_results
  }

@@

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun = FALSE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun = FALSE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun = FALSE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun = FALSE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun = FALSE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun = FALSE),
  list(name = "cem", method = "cem", estimand="ATT", rerun = FALSE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun = FALSE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun=FALSE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun=FALSE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun=FALSE),
  list(name = "twang_gbm", use="twanng", method = "twang", estimand="ATT", rerun = FALSE)
)

results <- run_matching_methods(
  data = m_data,
  formula = matchingFormula,
  methods = methods,
  rerun_all = FALSE,
  output_dir = ps_dir
)

>>>>

library(cobalt)
library(dplyr)
library(gee)
library(ggplot2)
library(kableExtra)
library(margins)
library(MatchIt)
library(purrr)
library(readr)
library(survey)
library(twang)
library(WeightIt)

@@

# Define function to run matching with disk storage and selective re-running
run_matching_methods <- function(data, formula, methods, rerun_model = FALSE, rerun_diagnostics = TRUE, output_dir = "ps_results") {
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }

@@

  for (params in methods) {
    filename <- file.path(output_dir, paste0(params$name, ".rds"))
    diagnostics_filename <- file.path(output_dir, paste0(params$name, "_diagnostics.rds"))

@@

    if (file.exists(filename) && !rerun_model) {
      print("Results are cached. Proceeding to next method...")
      method_results <- readRDS(filename)

      if (file.exists(diagnostics_filename) && !rerun_diagnostics) {
        diagnostics <- readRDS(diagnostics_filename)
        method_results <- c(method_results, diagnostics)
      }

      all_results[[params$name]] <- method_results
      next
    }

@@

    result <- tryCatch({
      if (params$method == "twang") {
        ps(formula = formula, data = data.frame(data), 
           n.trees = 10000, interaction.depth = 3,
           shrinkage = 0.01, estimand = params$estimand,
           stop.method = c("es.mean", "ks.max"))
      } else if (!is.null(params$use) && params$use == "weightit") {
        do.call(weightit, c(list(formula = formula, data = data), params))
      } else {
        do.call(matchit, c(list(formula = formula, data = data), params))
      }
    }, error = function(e) {
      message("Error in ", params$name, ": ", e)
      return(NULL)
    })

    if (is.null(result)) {
      next
    }

    balance_plot <- NULL
    common_support_plot <- NULL
    balance_summary <- NULL

    if (rerun_diagnostics) {
      balance_plot <- tryCatch({
        love.plot(result, stat = "mean.diffs", var.order = "adjusted", stars = "std")
      }, error = function(e) {
        message("Error generating balance plot for ", params$name, ": ", e)
        return(NULL)
      })

      if (!params$method %in% c("nearest_mahalanobis", "cem")) {
        common_support_plot <- tryCatch({
          if (params$method == "twang") {
            bal.plot(result, var.name = "prop.score.es.mean", which = "both")
          } else {
            bal.plot(result, var.name = "distance", which = "both")
          }
        }, error = function(e) {
          message("Error generating common support plot for ", params$name, ": ", e)
          return(NULL)
        })
      }

      balance_summary <- tryCatch({
        if (params$method == "twang") {
          summary(result)
        } else {
          bal.tab(result)
        }
      }, error = function(e) {
        message("Error generating balance summary for ", params$name, ": ", e)
        return(NULL)
      })
    }

    method_results <- list(result = result)
    saveRDS(method_results, file = filename)

    diagnostics <- list(balance_plot = balance_plot, common_support_plot = common_support_plot, summary = balance_summary)
    saveRDS(diagnostics, file = diagnostics_filename)

    all_results[[params$name]] <- c(method_results, diagnostics)
  }

@@

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "cem", method = "cem", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "twang_gbm", use="twanng", method = "twang", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE)
)

results <- run_matching_methods(
  data = m_data,
  formula = matchingFormula,
  methods = methods,
  rerun_model = FALSE,
  output_dir = ps_dir
)
